{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion}

### 2017-01-07 proposal/related.tex
do more justice to MC TM related work

Transactional memory (TM), first introduced in \cite{transactional-memory},
has received renewed attention in recent years with the announcement of Intel's Haswell architecture \cite{htm-haswell},
which promises to support hardware transactions using new x86 instructions.
Because hardware transactions can fail for any reason, not just on memory conflicts,
software TM remains relevant,
and recent works such as \cite{hybrid-htm-stm} and \cite{stm-relaxed-memory} enhance it to be nested with hardware transactions, or to be used on weak memory architectures, respectively;
which in turn produce ever more complicated concurrency models for people like me to tackle.
%
Testing approaches to TM are sparsely represented in the literature so far.
%The Weakest Reasonable Condition
\cite{specifying-verifying-tm} introduces several conditions to help specify/verify the correctness of TM implementations,
but not that of client programs which use TM.
{\em Learning from Mistakes} \cite{learning-from-mistakes} is a study of the characteristics of many different types of concurrency bugs;
it found that TM could potentially fix some, but not all, of the studied bugs,
although in some cases it must be combined with other concurrency primitives,
which indicates a major need for verifying transactional code with model checking.

>>>>

Transactional memory (TM), first introduced in \cite{transactional-memory},
has received renewed attention in recent years with the announcement of Intel's Haswell architecture \cite{htm-haswell},
which promises to support hardware transactions using new x86 instructions.
Because hardware transactions can fail for any reason, not just on memory conflicts,
software TM remains relevant,
and recent works such as \cite{hybrid-htm-stm} and \cite{stm-relaxed-memory} enhance it to be nested with hardware transactions, or to be used on weak memory architectures, respectively;
which in turn produce ever more complicated concurrency models for people like me to tackle.
%
Testing approaches for transactional programs are sparsely represented in the literature so far.
%The Weakest Reasonable Condition
Several related works \cite{tm-correctness,tm-completeness,specifying-verifying-tm} are building up to formal proofs of the correctness of TM {\em implementations},
but not that of client programs which use TM.
Getting closer, McRT STM \cite{mc-tm-with-spin} uses SPIN \cite{spin} to model check an STM implementation
up to 2 threads running 1 transaction each with up to 3 memory accesses.
This kind of verification is an important stepping stone for trusting the results Landslide will provide.
% TODO: make this point once you've added discussion of mutex_test to chap:quicksand.
% This kind of verification is analogous to the {\tt mutex\_test}
%
{\em Learning from Mistakes} \cite{learning-from-mistakes} studies the characteristics of many types of concurrency bugs;
it found that TM could potentially fix some, but not all, of the studied bugs,
although in some cases it must be combined with other concurrency primitives,
which indicates a major need for verifying transactional code with model checking.
